Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
data collection periods language refined

To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods. Two types of departures, namely non-linearity and heteroskedasticity, are considered with the corresponding data generating process being designed for data collection periods I and II. The data collection period III is designed for collecting human responses to null lineups such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to collect `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.

@@

```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used in data collection period I. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}

@@

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in data collection period I. This lineup is produced by the non-linearity model with $j = 6$ and uniform fitted values. The data plot location is $2^3 - 4$. All five subjects correctly identify the data plot from this lineup.

```{r different-n, fig.height = 2.67, fig.cap = "Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in data collection periods I, II and III."}

@@

```{r different-sigma, fig.height = 2, fig.cap = "For the non-lienearity model, the standard deviation of the error term $\\sigma$ affects the strength of the signal. Four different values of $\\sigma$ are used in the data collection period I."}

@@

```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in data collection period II. Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes."}

@@

```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in data collection period II to control the strength of the signal."}

@@

This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models. An experiment is conducted in three data collection periods with two simulation models designed for non-linearity and heteroskedasticity model defects being applied. Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allowed for multiple selections in a lineup protocol.

>>>>

To investigate the difference between conventional hypothesis testing and visual inference in the application of linear regression diagnostics, an experiment is conducted in three data collection periods. The two types of departures, non-linearity and heteroskedasticity, are primarily collected during data collection periods I and II. The data collection period III was designed primarily to measure human responses to null lineups so that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Additional lineups for both non-linearity and heteroskedasticity, using uniform $X$ distribution, were included so that the participants were evaluating some lineups with signal also. It would be too frustrating for participants to only be assigned lineups with all null plots. Overall, we collected `r vi_survey  %>% filter(!attention_check) %>% nrow()` evaluations on `r vi_survey  %>% filter(!attention_check) %>% pull(unique_lineup_id) %>% unique() %>% length()` unique lineups performed by `r vi_survey  %>% filter(!attention_check) %>% count(exp, set) %>% nrow()` subjects throughout three data collection periods.

@@

```{r different-shape-of-herimite, fig.height = 2, fig.cap = "Polynomial forms generated for the residual plots used to assess detecting non-linearity. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}

@@

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups used in non-linearity detection. This lineup is produced by the non-linearity model with $j = 6$ and uniform fitted values. The data plot location is $2^3 - 4$. All five subjects correctly identify the data plot from this lineup.

```{r different-n, fig.height = 2.67, fig.cap = "Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\\sigma = 1.5$, uniform $X$ distribution, and S shape. For these factor levels, only when $n=300$ is the S shape clearly visible."}

@@

```{r different-sigma, fig.height = 2, fig.cap = "Examining the effect of $\\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform $X$ distribution and the U shape. As $\\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\\sigma=4$."}

@@

```{r different-shape-of-heter, fig.height = 2.67, fig.cap = "Heteroskedasticity forms used in the experiment. Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively."}

@@

```{r different-b, fig.height = 2, fig.cap = "Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal. Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal."}

@@

This paper has demonstrated that conventional residual-based statistical tests are more sensitive to departures from assumptions for residuals than visual tests evaluated by humans in the diagnostics of linear regression models. An experiment is conducted in three data collection periods using two simulation models, one for for non-linearity and and the other for heteroskedasticity. Methods for obtaining the $p$-value of visual tests proposed by @vanderplas2021statistical are extended to allow for multiple selections in a lineup protocol.

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
data collection periods language refined

To investigate the difference between conventional hypothesis testing
and visual inference in the application of linear regression
diagnostics, an experiment is conducted in three data collection
periods. Two types of departures, namely non-linearity and
heteroskedasticity, are considered with the corresponding data
generating process being designed for data collection periods I and II.
The data collection period III is designed for collecting human
responses to null lineups such that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated. Overall, we plan to
collect 7974 evaluations on 1152 unique lineups performed by 443
subjects throughout three data collection periods.

@@

\caption{Polynomial forms generated for the residual plots used in data collection period I. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
\end{figure}

@@

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
used in data collection period I. This lineup is produced by the
non-linearity model with \(j = 6\) and uniform fitted values. The data
plot location is \(2^3 - 4\). All five subjects correctly identify the
data plot from this lineup.

@@

\caption{Number of observations $n$ have an impact on the strength of the signal. Three different values of $n$ are used in data collection periods I, II and III.}\label{fig:different-n}
\end{figure}

@@

\caption{For the non-lienearity model, the standard deviation of the error term $\sigma$ affects the strength of the signal. Four different values of $\sigma$ are used in the data collection period I.}\label{fig:different-sigma}
\end{figure}

@@

\caption{Heteroskedasticity forms used in data collection period II. Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle, butterfly and right-triangle shapes.}\label{fig:different-shape-of-heter}
\end{figure}

@@

\caption{Five different values of $b$ are used in data collection period II to control the strength of the signal.}\label{fig:different-b}
\end{figure}

@@

This paper has demonstrated that conventional residual-based statistical
tests are more sensitive to departures from assumptions for residuals
than visual tests evaluated by humans in the diagnostics of linear
regression models. An experiment is conducted in three data collection
periods with two simulation models designed for non-linearity and
heteroskedasticity model defects being applied. Methods for obtaining
the \(p\)-value of visual tests proposed by
\citet{vanderplas2021statistical} are extended to allowed for multiple
selections in a lineup protocol.

>>>>

To investigate the difference between conventional hypothesis testing
and visual inference in the application of linear regression
diagnostics, an experiment is conducted in three data collection
periods. The two types of departures, non-linearity and
heteroskedasticity, are primarily collected during data collection
periods I and II. The data collection period III was designed primarily
to measure human responses to null lineups so that the parameter
\(\alpha\) in Equation \ref{eq:pvalue-beta-binomial} can be estimated.
Additional lineups for both non-linearity and heteroskedasticity, using
uniform \(X\) distribution, were included so that the participants were
evaluating some lineups with signal also. It would be too frustrating
for participants to only be assigned lineups with all null plots.
Overall, we collected 7974 evaluations on 1152 unique lineups performed
by 443 subjects throughout three data collection periods.

@@

\caption{Polynomial forms generated for the residual plots used to assess detecting non-linearity. The four shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-shape-of-herimite}
\end{figure}

@@

Figure \ref{fig:example-poly-lineup} demonstrates one of the lineups
used in non-linearity detection. This lineup is produced by the
non-linearity model with \(j = 6\) and uniform fitted values. The data
plot location is \(2^3 - 4\). All five subjects correctly identify the
data plot from this lineup.

@@

\caption{Examining the effect of signal strength for the three different values of $n$ used in the experiment, for non-linear structure with fixed $\sigma = 1.5$, uniform $X$ distribution, and S shape. For these factor levels, only when $n=300$ is the S shape clearly visible.}\label{fig:different-n}
\end{figure}

@@

\caption{Examining the effect of $\sigma$ on the signal strength in the non-linearity detection, for $n=300$, uniform $X$ distribution and the U shape. As $\sigma$ increases the signal strength decreases, to the point that the U is almost unrecognisable when $\sigma=4$.}\label{fig:different-sigma}
\end{figure}

@@

\caption{Heteroskedasticity forms used in the experiment. Three different shapes ($a = -1, 0, 1$) are used in the experiment to create left-triangle, butterfly and right-triangle shapes, respectively.}\label{fig:different-shape-of-heter}
\end{figure}

@@

\caption{Five different values of $b$ are used in heteroskedasticity simulation to control the strength of the signal. Larger values of $b$ yield a bigger difference in variation, and stus stronger heteroskedasticity signal.}\label{fig:different-b}
\end{figure}

@@

This paper has demonstrated that conventional residual-based statistical
tests are more sensitive to departures from assumptions for residuals
than visual tests evaluated by humans in the diagnostics of linear
regression models. An experiment is conducted in three data collection
periods using two simulation models, one for for non-linearity and and
the other for heteroskedasticity. Methods for obtaining the \(p\)-value
of visual tests proposed by \citet{vanderplas2021statistical} are
extended to allow for multiple selections in a lineup protocol.
